The Ranked Nearest Neighbour Interchange space

### 2019-08-20 paper.tex
edit alg:max_dist_tree -- found counterexample for arbitrary trees

In \autocite{Gavryushkin2018-ol} an efficient algorithm for computing the $\rnni$ graph has been introduced.
We implemented this algorithm and could produce the graph for trees with up to $8$ taxa.
This enables us to compute exact distances between trees of this size.
It is in particular possible to compare the exact distances with distances we receive from algorithms we introduced previously and to verify the Cluster Theorem, which was raised in Section~\ref{section:split_theorem}, for small trees.
All algorithms have been implemented with Python and are available online. \todo{link?}

@@

Recall that the Section~\ref{section:diameter} revealed that the maximum distance computed by $\findpath$ for an arbitrary pair of trees equals the diameter of $\rnni$.
Thus, if the distance between two trees equals the diameter, $\findpath$ computes the correct distance.
Note that central for proving the convexity of the set of caterpillar trees in Section~\ref{section:caterpillar_convex} was Lemma~\ref{lemma:caterpillar_dist=diameter}.
We therefore aim to find a property for any pair of trees, not only caterpillar trees, with maximum distance $\Delta(\rnni) = \frac{(n-1)(n-2)}{2}$.
Following algorithm $\findpath$, we claim following conjecture:

@@

Provided Conjecture~\ref{conjecture:findpath_maxdist_correct} holds, there is a simple algorithm for computing a tree $R$ with maximum distance from any given tree $T$.
%TODO
%Explain this algorithm, not only Pseudo-Code!

\begin{algorithm}[H]
\caption{MAX\_DISTANCE\_TREE($T$)}
\label{alg:max_dist_tree}
\begin{algorithmic}[1]
	\STATE $R$ only contains cherry taxa $a, b$ of $T$ and has internal node with rank $n-1$
    \STATE In list $L$ all taxa $\{1,\ldots,n\}\setminus\{a,b\}$ are sorted such that the ranks of their parents in $T$ do not decrease
	\FOR {taxon $a$ in $L$}
		\STATE Add an internal node as parent of $a$ on any edge in $R$ such that the new internal node has rank one less than the previously lowest internal node
	\ENDFOR
	\RETURN $R$
\end{algorithmic}
\end{algorithm}

\begin{lemma}
    Let $T$ and $R$ be trees.
    Algorithm~\ref{alg:find_path} ($\findpath$) gives a path of length $\frac{(n-1)(n-2)}{2}$ between $T$ and $R$ if, and only if, $R$ can be generated from $T$ by applying Algorithm~\ref{alg:max_dist_tree}.
    \label{lemma:max_dist_lemmas_equivalence}
\end{lemma}

According to Lemma~\ref{lemma:max_dist_lemmas_equivalence}, Conjecture~\ref{conjecture:findpath_maxdist_correct} is equivalence to the conjecture that two trees have maximum distance if, and only if, one can be obtained from the other by applying Algorithm~\ref{alg:max_dist_tree}.

%TODO
%proof (short explanation)

If Conjecture~\ref{conjecture:findpath_maxdist_correct} is correct, it directly would follow that the radius of $\rnni$ is $rad(\rnni) = \Delta(\rnni) = \frac{(n-1)(n-2)}{2}$, where $rad(G) = \min\limits_u \max\limits_v d(u,v)$ for graph $G$ and nodes $u$ and $v$ in $G$.

>>>>

In \autocite{Gavryushkin2018-ol} an efficient algorithm for computing the $\rnni$ graph has been introduced.
We implemented this algorithm and could produce the graph for trees with up to $8$ taxa. \todo{Shall we say more about the algorithm here?}
This enables us to compute exact distances between trees of this size.
In particular, this enables us to compare exact distances with distances we receive from algorithms introduced in previous sections and to verify the Cluster Theorem, which was raised in Section~\ref{section:split_theorem}, for small trees.
All algorithms have been implemented with Python and are available online. \todo{link?}

@@

Recall that the Section~\ref{section:diameter} revealed that the maximum distance computed by $\findpath$ for an arbitrary pair of trees equals the diameter of $\rnni$.
Thus, if the distance between two trees equals the diameter, $\findpath$ computes the correct distance.
Following this observation, we claim following conjecture:

@@

Besides the diameter, the \emph{radius} of a graph $G$, which is defined by $rad(G) =  \min\limits_u \max\limits_v d(u,v)$ is a further interesting property of a graph.
In Conjecture~\ref{conjecture:radius} we claim that the radius of $\rnni$ equals its diameter.
Algorithm~\ref{alg:max_dist_tree} suggests this, since it computes for an arbitrary input tree $T$ a caterpillar tree $R$ with distance $d(T,R) = \frac{(n-1)(n-2)}{2}$ for $n \leq k$. \todo{$k = ?$, check computationally (or prove this -> see conjecture below)}

\begin{conjecture}
    It is $rad(\rnni) = \Delta(\rnni) = \frac{(n-1)(n-2)}{2}$.
    \label{conjecture:radius}
\end{conjecture}

%TODO
%Explain the following algorithm, not only Pseudo-Code!

\begin{algorithm}[H]
\caption{MAX\_DISTANCE\_TREE($T$)}
\label{alg:max_dist_tree}
\begin{algorithmic}[1]
	\STATE $R$ only contains cherry taxa $a, b$ of $T$ and has internal node with rank $n-1$
    \STATE In list $L$ all taxa $\{1,\ldots,n\}\setminus\{a,b\}$ are sorted such that the ranks of their parents in $T$ do not decrease
	\FOR {taxon $a$ in $L$}
		\STATE Add an internal node as parent of $a$ on an edge in $R$ such that the new internal node has rank one less than the previously lowest internal node and $R$ is a caterpillar tree
	\ENDFOR
	\RETURN $R$
\end{algorithmic}
\end{algorithm}

\begin{conjecture}
    Let $T$ be an arbitrary tree and $R$ a caterpillar tree.
    Algorithm~\ref{alg:find_path} ($\findpath$) gives a path of length $\frac{(n-1)(n-2)}{2}$ between $T$ and $R$ if, and only if, $R$ can be generated from $T$ by applying Algorithm~\ref{alg:max_dist_tree}.
    \label{conjecture:max_dist_caterpillar}
\end{conjecture}

%TODO
%prove this conjecture! (<= would be sufficient for the conjecture above!)
